\documentclass[12pt,a4paper]{article}

% ── packages ──────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{microtype}

% ── colours ───────────────────────────────────────────────────────────────────
\definecolor{snbblue}{HTML}{1F6FB2}
\definecolor{charcoal}{HTML}{2C2C2C}
\definecolor{midgray}{HTML}{888888}
\definecolor{lightgray}{HTML}{F5F5F5}

% ── hyperref ──────────────────────────────────────────────────────────────────
\hypersetup{
  colorlinks = true,
  linkcolor  = snbblue,
  urlcolor   = snbblue,
  citecolor  = snbblue,
  pdftitle   = {Swiss Confederation Debt Issuance: Instruments and Data},
  pdfauthor  = {SNB\_MoPo project},
}

% ── headers ───────────────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\rhead{\textcolor{midgray}{\small SNB\_MoPo project}}
\lhead{\textcolor{midgray}{\small Swiss Confederation Debt Issuance}}
\rfoot{\textcolor{midgray}{\small \thepage}}
\renewcommand{\headrulewidth}{0.4pt}

% ── section formatting ────────────────────────────────────────────────────────
\titleformat{\section}{\large\bfseries\color{charcoal}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries\color{snbblue}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\itshape\color{charcoal}}{}{0em}{}

% ── custom column types ───────────────────────────────────────────────────────
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

% ─────────────────────────────────────────────────────────────────────────────
\begin{document}
% ─────────────────────────────────────────────────────────────────────────────

\begin{center}
  {\LARGE\bfseries Swiss Confederation Debt Issuance}\\[0.4em]
  {\large Instruments, Practices, and Data}\\[1.2em]
  {\normalsize SNB\_MoPo project \quad|\quad
   \href{https://github.com/ThomasNellen/SNB_MoPo}{github.com/ThomasNellen/SNB\_MoPo}}\\[0.3em]
  {\small\textcolor{midgray}{Data vintage: April 2026 \quad|\quad
   Source: EFV / Federal Finance Administration}}
\end{center}

\vspace{0.5em}
\hrule height 1pt
\vspace{1.5em}

% ─────────────────────────────────────────────────────────────────────────────
\section{Institutional Framework}
% ─────────────────────────────────────────────────────────────────────────────

The \textbf{Federal Finance Administration} (EFV --- Eidgenössische Finanz\-verwaltung
/ Administration fédérale des finances) manages the Swiss Confederation's
financing needs.  It issues debt in two maturity segments:

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Money market}: Geldmarktbuchforderungen (GMBF), maturities up
    to one year.
  \item \textbf{Capital market}: Confederation bonds (Eidgenössische Anleihen),
    maturities from approximately 4 to 50 years.
\end{itemize}

\subsection{The SNB as Fiscal Agent}

The Swiss National Bank (SNB) acts as the Confederation's fiscal agent and
\emph{bank}.  In this role it:

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item Manages the Confederation's sight-deposit account at the SNB (the
    federal treasury account).
  \item Operates the auction platform for both GMBF and bond issuances on
    behalf of the EFV.
  \item Conducts settlement of issuance proceeds and redemption payments.
  \item Holds and administers the Confederation's \emph{own holdings}
    (Eigenbestand) --- tranches retained at issuance and placed opportunistically
    at a later date.
\end{itemize}

\subsection{The Auction Platform}

All issuances take place on the \textbf{SIX Repo Ltd} platform (formerly the
Swiss repo system operated by SIX), using a uniform electronic tender
process.  Eligible participants are banks and securities dealers domiciled in
Switzerland, and --- for bond auctions --- Swiss insurance companies.  The SNB
controls access and operates the technical infrastructure; EFV sets the
issuance parameters (volume, maturities, minimum price or yield).

\medskip
\noindent The same platform hosts the SNB's own monetary-policy repo operations
(see the companion document \texttt{snb\_mopo\_overview.pdf}).  Both the
Confederation's debt auctions and the SNB's daily liquidity operations are
conducted in this single electronic marketplace.

% ─────────────────────────────────────────────────────────────────────────────
\section{Geldmarktbuchforderungen (GMBF)}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{What They Are}

GMBF (literally ``money-market book-entry claims'') are short-term, \emph{zero-coupon}
debt instruments issued by the Swiss Confederation at a discount and redeemed
at face value (CHF\,100) on the maturity date.  They are
book-entry securities (no physical certificate) held in the SIX Securities
Services central depository, and they trade in the secondary market, though
liquidity there is thin.

\subsection{Maturities and Target Stock}

Three standard tenors are in regular use:

\begin{center}
\begin{tabular}{cccc}
  \toprule
  Tenor & Approximate days & Typical ISIN count & Rotation \\
  \midrule
  3-month  & $\sim$91 days  & 11 outstanding & Weekly \\
  6-month  & $\sim$182 days &  3 outstanding & Monthly (roughly) \\
  12-month & $\sim$364 days &  2 outstanding & Bimonthly (roughly) \\
  \bottomrule
\end{tabular}
\end{center}

\medskip
\noindent The EFV targets a total of \textbf{16 GMBF series} outstanding at any
one time (11\,+\,3\,+\,2).  The target outstanding volume is set annually; in
2025 it was CHF\,8--14\,billion, with actual year-end stock of CHF\,8.4\,bn.

\subsection{Auction Mechanics}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Frequency}: every \textbf{Tuesday} (with rare exceptions for
    public holidays, when the auction shifts to Wednesday).
  \item \textbf{Subscription window}: \textbf{10:00--10:30 CET/CEST} on the
    SIX Repo Ltd platform.
  \item \textbf{Auction type}: rate tender with single-price (Dutch) allocation
    --- all successful bidders receive the same marginal price.  Non-competitive
    (price-less) bids are accepted up to a cap.
  \item \textbf{Settlement}: \textbf{T+2} (Thursday value date in the standard
    case).
  \item \textbf{Issuance at discount}: each auction produces a \emph{new} ISIN
    for that specific maturity date, priced below or around par.  A price above
    100 (premium) implies a negative yield.
  \item \textbf{Results publication}: the EFV publishes full results
    (price, yield, bids, allotment) on its website immediately after the
    auction; the annual XLSX is updated weekly.
\end{itemize}

\subsection{Pricing and Yields}

GMBF yields closely track the SNB policy rate at the corresponding tenor.
During the negative-rate era (2015--2022) GMBF yields were consistently
negative, with the 3-month yield typically around $-0.75$\% p.a.\
(matching the SNB policy rate); from mid-2022 to end-2024 they turned
positive, peaking around $+1.8$\% p.a.; they have returned to slightly
negative territory in 2025--2026 following the SNB's rate cuts.

The bid-to-cover ratio on the 3-month tranche is typically \textbf{3--6$\times$},
reflecting strong institutional demand.  The 6- and 12-month tranches
attract somewhat lower cover ratios.

% ─────────────────────────────────────────────────────────────────────────────
\section{Confederation Bonds (Eidgenössische Anleihen)}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{What They Are}

Confederation bonds are long-term, fixed-rate bullet bonds issued by the
Swiss Confederation.  They pay an annual coupon and are redeemed at par at
maturity.  Maturities range from approximately \textbf{4 to 50 years}, making
Switzerland one of the few sovereign issuers with a regularly issued
ultra-long segment.  Outstanding volume across all series totals approximately
CHF\,75\,billion placed on the market (as of Q1\,2026), with 23 live series.

\subsection{Auction Mechanics}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Frequency}: \textbf{monthly}, on the \textbf{second Wednesday}
    of each month.  August is excluded; October and December auctions are
    optional depending on financing needs.  This gives 9--11 auctions per year.
  \item \textbf{Subscription window}: \textbf{10:30--11:00 CET/CEST} on the
    SIX Repo Ltd platform.
  \item \textbf{Two tranches per auction}: each auction reopens exactly two
    existing bond series (tap issuances / Aufstockungen).  This concentrates
    liquidity in a small number of benchmark series rather than creating many
    small illiquid lines.
  \item \textbf{Auction type}: price tender with uniform (Dutch) pricing ---
    all accepted bidders pay the marginal price.  Non-competitive bids are
    accepted.
  \item \textbf{Settlement}: approximately \textbf{T+14} (the second Friday
    after the auction date, or thereabouts).
  \item \textbf{Own holdings (Eigenbestand)}: the EFV may retain a portion of
    the allotment for its own account (column M in the results file).  These
    own-holdings tranches are not counted as ``placed on market'' and are
    subsequently sold bilaterally at prevailing market prices when demand
    arises (see Section~\ref{sec:own}).
  \item \textbf{Results publication}: the EFV releases results immediately
    after the auction.  The annual XLSX is updated after each monthly auction.
\end{itemize}

\subsection{Series Management and Reopenings}

Rather than issuing a new ISIN at each auction, the EFV builds up each
series progressively through multiple tap issuances over several years until
the desired benchmark size is reached (typically CHF\,3--6\,billion placed
on market).  A new series is started (identified by a provisional ISIN that
becomes definitive at first settlement) only occasionally, when the maturity
profile requires filling a gap.

Bond names follow the convention \textbf{``Eidg.\,DD.MM.YY/YY''}, where the
first date is the first settlement date and the last two digits are the
maturity year (e.g.\ ``Eidg.\,23.06.21/35'' = Confederation bond first
settled 23 June 2021, maturing 2035).

\subsection{Own Holdings and Bilateral Sales}
\label{sec:own}

At each auction the EFV may choose not to place the full allotment with market
participants, retaining a portion as \emph{own holdings}.  These are recorded
in the auction results as ``Additional own holdings not yet placed.''  The
retained volume is subsequently sold bilaterally at market prices, typically
at the request of a specific counterparty.  These transactions are not
separately announced; they are disclosed only in the EFV's annual activity
report as aggregate figures (e.g.\ CHF\,138\,mn across 4 transactions in 2025).

This mechanism is functionally similar to a private placement in that it
involves a direct transaction between the EFV and a single counterparty at
a negotiated price, but it uses existing (already-issued) ISINs rather than
creating new ones.

\subsection{Yield Levels}

Swiss Confederation bonds trade at among the lowest yields in the world,
reflecting the AAA-rated sovereign's safe-haven status and the structural
CHF demand.  Yields have fluctuated substantially over the dataset period:

\begin{center}
\begin{tabular}{lcc}
  \toprule
  Period & 10-year range (approx.) & Dominant driver \\
  \midrule
  2011--2014 & 0.5\%--1.5\% & Euro-area crisis / safe-haven flows \\
  2015--2021 & $-0.5$\%--0\%  & SNB negative rates, QE spillovers \\
  2022--2023 & 0\%--1.5\%   & Global rate normalisation \\
  2024--2026 & 0.2\%--0.8\% & SNB easing cycle \\
  \bottomrule
\end{tabular}
\end{center}

% ─────────────────────────────────────────────────────────────────────────────
\section{Relationship to SNB Monetary Policy}
% ─────────────────────────────────────────────────────────────────────────────

Several interactions between Confederation debt issuance and SNB operations
are worth noting:

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Settlement flows}: GMBF settlement (Thursdays) and SNB Bill
    settlement (also approximately T+4 from Thursdays) both hit the
    interbank CHF market on similar dates, potentially creating temporary
    liquidity effects that show up in SARON.
  \item \textbf{Collateral}: Confederation bonds are the primary high-quality
    collateral used in Swiss repo markets (GC basket), meaning they
    underpin both the SNB's monetary-policy repos and private-sector
    short-term funding.  New issuances expand the collateral pool available
    to the banking system.
  \item \textbf{Yield curve}: the SNB's policy rate anchors the short end;
    Confederation bond auctions reveal the long end.  The slope reflects
    market expectations about future SNB rates and the term premium.
  \item \textbf{Coordination}: while EFV and SNB act independently, the
    publication of the annual GMBF and bond issuance calendars allows
    money-market participants to anticipate supply and manage their
    portfolio duration.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Variable Definitions}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{GMBF Auction Results (\texttt{efv\_gmbf\_auctions.csv})}

746 auction rows covering January 2012 to April 2026 (one row per auction per
ISIN).  Yields stored in per cent per annum; amounts in CHF millions.

\begin{longtable}{L{4.0cm} L{2.3cm} L{7.5cm}}
  \toprule
  \textbf{Column} & \textbf{Type} & \textbf{Description} \\
  \midrule
  \endhead
  \bottomrule
  \endfoot
  \texttt{auction\_date}         & date  & Date of the auction (Tuesday) \\[3pt]
  \texttt{settlement\_date}      & date  & Value date, T+2 (Thursday) \\[3pt]
  \texttt{maturity\_date}        & date  & Redemption date at par \\[3pt]
  \texttt{term\_days}            & int   & Exact calendar days from settlement
                                           to maturity \\[3pt]
  \texttt{term\_bucket}          & str   & Rounded tenor: \texttt{91d}
                                           ($\leq$100\,d), \texttt{182d}
                                           ($\leq$200\,d), \texttt{364d}
                                           (otherwise) \\[3pt]
  \texttt{series}                & float & Legacy series number (2012--2025
                                           only; \texttt{NaN} from 2026
                                           when dropped by EFV) \\[3pt]
  \texttt{isin}                  & str   & ISIN of the specific maturity
                                           (CH\ldots) \\[3pt]
  \texttt{total\_bids\_chf\_mn}  & float & Aggregate bids submitted
                                           (CHF mn) \\[3pt]
  \texttt{bids\_no\_price\_chf\_mn} & float & Non-competitive bids (without
                                           price indication; CHF mn);
                                           always fully accepted \\[3pt]
  \texttt{issue\_volume\_chf\_mn}& float & Amount allotted to market
                                           (CHF mn) \\[3pt]
  \texttt{bid\_cover}            & float & Bid-to-cover ratio
                                           $= \text{total bids} /
                                           \text{issue volume}$;
                                           derived \\[3pt]
  \texttt{price}                 & float & Single marginal price
                                           (\% of face value; e.g.\
                                           100.019 or 99.912) \\[3pt]
  \texttt{yield\_pct}            & float & Annualised yield at the marginal
                                           price (\% p.a.); negative when
                                           price $>$ 100 \\
\end{longtable}

\subsection{Bond Auction Results (\texttt{efv\_bond\_auctions.csv})}

297 tranche rows across 167 auction dates, covering November 2010 to April
2026 (two rows per auction date, one per reopened series).

\begin{longtable}{L{4.4cm} L{2.3cm} L{7.1cm}}
  \toprule
  \textbf{Column} & \textbf{Type} & \textbf{Description} \\
  \midrule
  \endhead
  \bottomrule
  \endfoot
  \texttt{auction\_date}          & date  & Date of the auction (2nd Wednesday
                                            of the month) \\[3pt]
  \texttt{settlement\_date}       & date  & Value date ($\approx$T+14) \\[3pt]
  \texttt{maturity\_date}         & date  & Final maturity of the bond \\[3pt]
  \texttt{residual\_maturity\_yrs}& float & Years from auction date to maturity;
                                            derived \\[3pt]
  \texttt{prov\_isin}             & str   & Provisional ISIN (becomes the
                                            definitive ISIN at settlement;
                                            ``\texttt{-}'' for new series in
                                            some recent rows) \\[3pt]
  \texttt{fungible\_isin}         & str   & ISIN of the existing benchmark
                                            series this tranche is fungible
                                            with (i.e.\ same bond, different
                                            settlement) \\[3pt]
  \texttt{bond\_name}             & str   & EFV bond label, format
                                            ``Eidg.\,DD.MM.YY/YY'' \\[3pt]
  \texttt{coupon\_pct}            & float & Annual coupon rate (\% p.a.;
                                            0 = zero-coupon bond) \\[3pt]
  \texttt{total\_bids\_chf\_mn}   & float & Aggregate bids (CHF mn) \\[3pt]
  \texttt{bids\_no\_price\_chf\_mn} & float & Non-competitive bids (CHF mn) \\[3pt]
  \texttt{issue\_volume\_chf\_mn} & float & Amount placed on market (CHF mn);
                                            excludes own holdings \\[3pt]
  \texttt{bid\_cover}             & float & Bid-to-cover ratio; derived \\[3pt]
  \texttt{price}                  & float & Marginal price (\% of face value) \\[3pt]
  \texttt{yield\_pct}             & float & Yield to maturity at marginal price
                                            (\% p.a.) \\[3pt]
  \texttt{own\_holdings\_not\_placed\_chf\_mn} & float & Volume retained by the
                                            Confederation as own holdings
                                            at this auction (CHF mn);
                                            sold bilaterally later \\
\end{longtable}

\subsection{Outstanding Bonds Snapshot (\texttt{efv\_bonds\_outstanding.csv})}

One row per live bond series; snapshot dated 31 March 2026.  23 series with a
combined placed volume of CHF\,75.4\,bn and CHF\,6.9\,bn own holdings still
available for bilateral sale.

\begin{center}
\begin{tabular}{L{5.0cm} L{2.2cm} L{6.6cm}}
  \toprule
  \textbf{Column} & \textbf{Type} & \textbf{Description} \\
  \midrule
  \texttt{isin} & str & Definitive ISIN (CH\ldots) \\[3pt]
  \texttt{bond\_name} & str & EFV bond label \\[3pt]
  \texttt{maturity\_date} & date & Final maturity \\[3pt]
  \texttt{coupon\_pct} & float & Annual coupon (\%) \\[3pt]
  \texttt{total\_issued\_incl\_own\_chf\_mn} & float &
    Total volume issued including all own holdings ever issued (CHF mn) \\[3pt]
  \texttt{total\_placed\_on\_market\_chf\_mn} & float &
    Volume placed with market participants (excl.\ own holdings; CHF mn) \\[3pt]
  \texttt{own\_holdings\_placed\_chf\_mn} & float &
    Own holdings that have been placed bilaterally since first issuance
    (CHF mn) \\[3pt]
  \texttt{own\_holdings\_available\_chf\_mn} & float &
    Own holdings still held by the Confederation and available for future
    bilateral sale (CHF mn) \\[3pt]
  \texttt{snapshot\_date} & date & Reference date of the snapshot \\
  \bottomrule
\end{tabular}
\end{center}

% ─────────────────────────────────────────────────────────────────────────────
\section{Data Sources and Reproducibility}
% ─────────────────────────────────────────────────────────────────────────────

All data are hosted by the EFV at
\url{https://www.efv.admin.ch/efv/en/home/finanzberichterstattung/}.
The SNB website (\url{https://www.snb.ch/en/the-snb/mandates-goals/services-to-the-confederation})
links to the same EFV files and publishes individual press releases for each
bond auction.  \textbf{No API or CSV endpoint exists}; all machine-readable data
is in XLSX format only.

\begin{center}
\begin{tabular}{L{3.5cm} L{10.5cm}}
  \toprule
  \textbf{Dataset} & \textbf{Direct download URL} \\
  \midrule
  GMBF auction results &
    \url{https://www.efv.admin.ch/dam/en/sd-web/d9FElh8dV0db/resultate-gmbf.xlsx} \\[4pt]
  Bond auction results &
    \url{https://www.efv.admin.ch/dam/en/sd-web/AFFQnqhWiEVW/resultate-anleihen.xlsx} \\[4pt]
  Outstanding bonds &
    \url{https://www.efv.admin.ch/dam/en/sd-web/K9gjjBZzAoPz/ausstehende-anleihen.xlsx} \\[4pt]
  GMBF issuance calendar &
    \url{https://www.efv.admin.ch/dam/en/sd-web/962GRmuHDjrK/GMBF-Kalender-2026.pdf} \\[4pt]
  Bond issuance calendar &
    \url{https://www.efv.admin.ch/dam/en/sd-web/kb0UBOFNnfjG/Kalender-Anleihen-2026.pdf} \\
  \bottomrule
\end{tabular}
\end{center}

\medskip
\noindent The download and parsing script is at
\texttt{code/efv\_debt\_analysis.py} in the project repository.  Running it
from the project root reproduces all three processed datasets from scratch.
Raw XLSX files are cached in \texttt{input/raw/} (gitignored) and only
re-downloaded if absent.

\medskip
\noindent\textbf{Update cadence:}
\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item GMBF file: updated every Tuesday after the auction.
  \item Bond file: updated after each monthly auction (second Wednesday).
  \item Outstanding bonds: updated by EFV periodically; typically reflects
    end of previous quarter.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\end{document}
